Measurement of nuclear excitation functions for proton induced reactions (E\texorpdfstring{$_{\text{p}}$\,=\,}{Ep = 40--90}40--90 MeV) on natural Nb

### 2018-04-05 nb_px_paper/nb_body_text.tex
Fixed Nb(p,x) fit of appendix tables and figures.  Fixed usage of micro symbol to avoid breaking typsetting.

Accurate cross section measurements using  activation methods benefit from well-characterized monitor reactions.  
Currently there is 
% , which  currently
a paucity of such data at intermediate energies, and much of what exists have  high uncertainties (\textgreater15\%).
% One significant obstacle to both high-fidelity measurements of production cross sections for emerging medical radionuclides, as well as conventional isotope production, is a lack of well-characterized  monitor reaction standards.  
% This is particularly true for higher-energy charged particle beams, where there is currently a paucity of such data. 
% Lee - cut?
Indeed, the development of new monitor reaction standards and the improved evaluation of existing standards is one of the areas of greatest cross-cutting need for nuclear data \cite{bernstein2015nuclear}. 
% Charged particle monitor reaction data currently exists for several low-to-intermediate energy charged particle beams (E \textless\ 50 MeV / A), but even the experimental data used for this evaluation are  sparse above approximately 30 MeV / A and  uncertainties in experimental cross sections are large (6--15\%) \cite{gul2001charged}. 
% While work is needed to improve upon existing monitor reaction data, the development of n
% Lee - cut?
New reactions can expand the available range of options for the monitoring of charged particle beams.
This work 
% in particular seeks to improve this range of options by 
is an attempt to characterize a new monitor reaction for  proton beams in excess of 40 MeV, for possible use at isotope production facilities such as the  Brookhaven Linac Isotope Producer (BLIP) at Brookhaven National Laboratory, 
% IPF
the Isotope Production Facility (IPF) at  Los Alamos National Laboratory, or the Separated Sector Cyclotron at the 
% iThemba LABS
iThemba Laboratory for Accelerator Based Sciences.
% \comment{Stephen: I'm having a hard time following this paragraph. Some of your syntax overlaps significantly with nuclear medicine \& medical physics terminology with different intended meanings. For example 'dosimetry' is typically taken to mean the measurement of absorbed energy per unit mass (J/kg) and associated measurements/specifications. Another example, 'dosimetry standard' is used quite often when referring to NIST air kerma reference standards which are the bedrock of Accredited Dosimetry Calibration Lab (ADCL) measurements. As you are more familiar with classic Nuclear Physics writing, perhaps you better know how this language will be interpreted. }

@@

One  reaction which satisfies these requirements is that of a new, intermediate-energy proton monitor reaction standard based on \ce{^{93}Nb}(p,4n)\ce{^{90}Mo}. 
Niobium is naturally mono-isotopic, readily  available commercially in high purity, is fairly chemically inert, and can easily be rolled down to foils as thin as 1 $\micro$m.  
\ce{^{90}Mo} also has 
% excellent decay properties ---
a sufficiently long lifetime ($\epsilon=100\%, t_{1/2}=5.56 \pm 0.09$ h \cite{Browne1997})  
% allows it to be counted at leisure without fear of the product \ce{^{90}Mo} decaying away excessively between end-of-bombardment and the start of counting, and it possesses
and seven strong, distinct gamma lines (notably its 122.370 keV [$I_\gamma = 64 \pm 3\%$] and 257.34 keV [$I_\gamma = 78 \pm 4\%$] lines) which can be used to uniquely and easily   quantify \ce{^{90}Mo} production. 
In addition,  \ce{^{90}Mo} is completely immune from (n,x) production on  \ce{^{93}Nb}, being produced only via the primary proton beam, and the \ce{^{90}Mo} decay lines can only be observed in its decay, as its daughter, \ce{^{90}Nb}, is also unstable and decays via $\epsilon$ to stable \ce{^{90}Zr}.

@@

A stacked-target design was utilized for this work in order that the (p,x) cross sections for each reaction channel could be measured at multiple energy positions in a single irradiation \cite{Cumming1963}. 
% For targets, a
A series of nominal 25 \micro m \ce{^{nat}Nb} foils (99.8\%, lot \#T23A035), 25 \micro m \ce{^{nat}Al} foils (99.999\%, lot \#M06C032), and 50 \micro m \ce{^{nat}Cu} foils (99.9999\%, lot \#N26B062) were used (all from Alfa Aesar, Ward Hill, MA, 01835, USA) targets were used.
Six foils of each metal were cut down to 2.5 x 2.5 cm squares and characterized --- for each foil, length and width measurements were taken at four different locations using a digital caliper (Mitutoyo America Corp.), thickness measurements were taken at four different locations using a digital micrometer (Mitutoyo America Corp.), and four mass measurements were taken using an analytical balance after cleaning the foils with isopropyl alcohol.
Using these length, width, and mass readings, the areal density and its uncertainty (in mg/cm$^2$) for each foil was calculated.
% , along with the propagated uncertainty in areal density.
The foils were tightly sealed into \enquote{packets} using two pieces of  3M 5413-Series Kapton polyimide film tape --- each piece of tape consists of 43.2 \micro m of a silicone adhesive (nominal 4.79 mg/cm$^2$) on 25.4 \micro m of a polyimide backing (nominal 3.61 mg/cm$^2$).
The sealed foils were mounted over the hollow center of a 1.575 mm-thick plastic frame.
One \ce{^{nat}Al}, one \ce{^{nat}Cu}, and one \ce{^{nat}Nb} mounted foil were bundled together using baling wire for each energy position.
These foil packet bundles were lowered into the beamline by inserting them into a  water-cooled production target box.
% These six foil packet bundles were inserted into one of six  
The box, seen in \autoref{fig:target_stack}, is machined from 6061 aluminum alloy, has a thin (0.64 mm) Inconel beam entrance window, and  contains 6 \enquote{energy positions} for targets, formed by  5 slabs of 6061 aluminum alloy (previously characterized) which serve as proton energy degraders  between energy positions.
% At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor --- after end-of-bombardment (EoB), decay radiation emitted from these activated stainless steel foils were used to develop radiochromic film (Gafchromic EBT3), revealing the spatial profile of the beam entering and exiting the stack.
After loading all targets in the stack, the lid of the target box is sealed in place, using an inset o-ring to create a water-tight seal, and the box is lowered through a hot cell into the beamline, where it sits electrically isolated.
The specifications of the target stack design for this work is presented in \autoref{tab:stack_table}.

This target stack was assembled and irradiated at the Isotope Production Facility (IPF) at the Los Alamos National Laboratory (LANL), using the LANSCE linear accelerator. 
The stack was irradiated for approximately 2 hours with a nominal current of 1 mA, using a 50 \micro s pulse at a frequency of 2 Hz, for an anticipated integral current of 205.9 nAh.
The beam current, measured using an inductive pickup, remained stable under these conditions for the duration of the irradiation, with the exception of approximately 70\,s of downtime, which occurred approximately 3 minutes into irradiation.
The proton beam incident upon the stack's Inconel beam entrance window had an average energy of 100 MeV determined via time-of-flight, with an approximately Gaussian energy distribution width of 0.1 MeV --- this energy profile was used for all later analysis.
At the end of the irradiation, the target stack was withdrawn from the beamline into the IPF hot cell, where it was disassembled and the activated foils removed using robotic manipulators.
The activated foils were cleaned of all surface contamination, and transported to a counting lab for gamma spectrometry, which started approximately 6 hours following end-of-bombardment.

@@

\caption{Specifications of the  target stack design in the present work. The proton beam enters the stack upstream of the 249.8 \micro m SS profile monitor, and is transported through the stack in the order presented here. The 6061 aluminum degraders have a measured density of approximately 2.80 g/cm$^3$. Their areal densities were determined using the variance minimization techniques described  in this work  and the earlier paper by Graves \etal\ \cite{Graves2016}. At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor --- after end-of-bombardment (EoB), decay radiation emitted from these activated stainless steel foils were used to develop radiochromic film (Gafchromic EBT3), revealing the spatial profile of the beam entering and exiting the stack.}

@@

SS profile monitor & 249.8 \micro m         & 194.56                                      & 0.29                      \\
Al-1               & 25.0 \micro m          & 6.52                                        & 0.72                      \\
Cu-1               & 61.3 \micro m          & 53.74                                       & 0.15                      \\
Nb-1               & 30.0 \micro m          & 23.21                                       & 0.17                      \\

@@

Al-2               & 25.5 \micro m          & 6.48                                        & 0.36                      \\
Cu-2               & 61.8 \micro m          & 53.85                                       & 0.17                      \\
Nb-2               & 30.8 \micro m          & 22.91                                       & 0.17                      \\

@@

Al-3               & 25.8 \micro m          & 6.47                                        & 0.31                      \\
Cu-3               & 61.5 \micro m          & 53.98                                       & 0.11                      \\
Nb-3               & 31.0 \micro m          & 22.91                                       & 0.24                      \\

@@

Al-4               & 26.3 \micro m          & 6.51                                        & 0.41                      \\
Cu-4               & 61.3 \micro m          & 53.46                                       & 0.22                      \\
Nb-4               & 30.8 \micro m          & 22.55                                       & 0.25                      \\

@@

Al-5               & 26.5 \micro m          & 6.48                                        & 0.29                      \\
Cu-5               & 61.5 \micro m          & 53.57                                       & 0.11                      \\
Nb-5               & 30.8 \micro m          & 22.11                                       & 0.25                      \\

@@

Al-6               & 26.3 \micro m          & 6.48                                        & 0.62                      \\
Cu-6               & 62.0 \micro m          & 53.84                                       & 0.32                      \\
Nb-6               & 31.3 \micro m          & 22.12                                       & 0.13                      \\
SS profile monitor & 124.4 \micro m         & 101.34                                      & 0.23                      \\ \bottomrule

@@

@@

@@

Notably, this work serves as the most well-characterized measurement of the \ce{^{nat}Nb}(p,x)\ce{^{90}Mo} reaction below 100 MeV to date, with cross sections measured  at the 4--6\% uncertainty level.
This is important, as it presents the first step towards characterizing this reaction for use as a proton monitor reaction standard below 100 MeV.
% The quantification of \ce{^{90}Mo} activity proved to be one of the most straightforward of all reaction products observed in this work.
\ce{^{nat}Nb}(p,x)\ce{^{90}Mo} can only be populated through the (p,4n) reaction channel, so no corrections for (n,x) contamination channels or decay down the A=90 isobar are needed.
% , which makes activity quantification trivial.
% In addition, 
\ce{^{90}Mo}  possesses seven strong, distinct gamma lines which can easily  be used for its identification and quantification.
Finally, the production of \ce{^{90}Mo}  in the 40--90 MeV region is quite strong, with a peak cross section of approximately 120 mb.
Combining the reaction yield and gamma abundance, the use of approximately 23 mg/cm$^2$ Nb targets easily provided sufficient counting statistics for activity quantification in the 40--90 MeV region.
This result presents the first step towards the use of \ce{^{90}Mo} as a clean and precise   charged particle monitor reaction standard in irradiations up to approximately 24 hours in duration.

>>>>

Accurate cross section measurements using  activation methods benefit from well- characterized monitor reactions.  
Currently there is 
% , which  currently
a paucity of such data at intermediate energies, and much of what exists have  high uncertainties (\textgreater15\%).
% One significant obstacle to both high-fidelity measurements of production cross sections for emerging medical radionuclides, as well as conventional isotope production, is a lack of well-characterized  monitor reaction standards.  
% This is particularly true for higher-energy charged particle beams, where there is currently a paucity of such data. 
% Lee - cut?
Indeed, the development of new monitor reaction standards and the improved evaluation of existing standards is one of the areas of greatest cross-cutting need for nuclear data \cite{bernstein2015nuclear}. 
% Charged particle monitor reaction data currently exists for several low-to-intermediate energy charged particle beams (E \textless\ 50 MeV / A), but even the experimental data used for this evaluation are  sparse above approximately 30 MeV / A and  uncertainties in experimental cross sections are large (6--15\%) \cite{gul2001charged}. 
% While work is needed to improve upon existing monitor reaction data, the development of n
% Lee - cut?
New reactions can expand the available range of options for the monitoring of charged particle beams.
This work 
% in particular seeks to improve this range of options by 
is an attempt to characterize a new monitor reaction for  proton beams in excess of 40 MeV, for possible use at isotope production facilities such as the  Brookhaven Linac Isotope Producer (BLIP) at Brookhaven National Laboratory, 
% IPF
the Isotope Production Facility (IPF) at  Los Alamos National Laboratory, or the Separated Sector Cyclotron at the 
% iThemba LABS
iThemba Laboratory for Accelerator Based Sciences.
% \comment{Stephen: I'm having a hard time following this paragraph. Some of your syntax overlaps significantly with nuclear medicine \& medical physics terminology with different intended meanings. For example 'dosimetry' is typically taken to mean the measurement of absorbed energy per unit mass (J/kg) and associated measurements/specifications. Another example, 'dosimetry standard' is used quite often when referring to NIST air kerma reference standards which are the bedrock of Accredited Dosimetry Calibration Lab (ADCL) measurements. As you are more familiar with classic Nuclear Physics writing, perhaps you better know how this language will be interpreted. }

@@

One  reaction which satisfies these requirements is that of a new, intermediate-energy proton monitor reaction standard based on \ce{^{93}Nb}(p,4n)\ce{^{90}Mo}. 
Niobium is naturally mono-isotopic, readily  available commercially in high purity, is fairly chemically inert, and can easily be rolled down to foils as thin as 1 \mmicro m.  
\ce{^{90}Mo} also has 
% excellent decay properties ---
a sufficiently long lifetime ($\epsilon=100\%, t_{1/2}=5.56 \pm 0.09$ h \cite{Browne1997})  
% allows it to be counted at leisure without fear of the product \ce{^{90}Mo} decaying away excessively between end-of-bombardment and the start of counting, and it possesses
and seven strong, distinct gamma lines (notably its 122.370 keV [$I_\gamma = 64 \pm 3\%$] and 257.34 keV [$I_\gamma = 78 \pm 4\%$] lines) which can be used to uniquely and easily   quantify \ce{^{90}Mo} production. 
In addition,  \ce{^{90}Mo} is completely immune from (n,x) production on  \ce{^{93}Nb}, being produced only via the primary proton beam, and the \ce{^{90}Mo} decay lines can only be observed in its decay, as its daughter, \ce{^{90}Nb}, is also unstable and decays via $\epsilon$ to stable \ce{^{90}Zr}.

@@

A stacked-target design was utilized for this work in order that the (p,x) cross sections for each reaction channel could be measured at multiple energy positions in a single irradiation \cite{Cumming1963}. 
% For targets, a
A series of nominal 25 \mmicro m \ce{^{nat}Nb} foils (99.8\%, lot \#T23A035), 25 \mmicro m \ce{^{nat}Al} foils (99.999\%, lot \#M06C032), and 50 \mmicro m \ce{^{nat}Cu} foils (99.9999\%, lot \#N26B062) were used (all from Alfa Aesar, Ward Hill, MA, 01835, USA) targets were used.
Six foils of each metal were cut down to 2.5 x 2.5 cm squares and characterized --- for each foil, length and width measurements were taken at four different locations using a digital caliper (Mitutoyo America Corp.), thickness measurements were taken at four different locations using a digital micrometer (Mitutoyo America Corp.), and four mass measurements were taken using an analytical balance after cleaning the foils with isopropyl alcohol.
Using these length, width, and mass readings, the areal density and its uncertainty (in mg/cm$^2$) for each foil was calculated.
% , along with the propagated uncertainty in areal density.
The foils were tightly sealed into \enquote{packets} using two pieces of  3M 5413-Series Kapton polyimide film tape --- each piece of tape consists of 43.2 \mmicro m of a silicone adhesive (nominal 4.79 mg/cm$^2$) on 25.4 \mmicro m of a polyimide backing (nominal 3.61 mg/cm$^2$).
The sealed foils were mounted over the hollow center of a 1.575 mm-thick plastic frame.
One \ce{^{nat}Al}, one \ce{^{nat}Cu}, and one \ce{^{nat}Nb} mounted foil were bundled together using baling wire for each energy position.
These foil packet bundles were lowered into the beamline by inserting them into a  water-cooled production target box.
% These six foil packet bundles were inserted into one of six  
The box, seen in \autoref{fig:target_stack}, is machined from 6061 aluminum alloy, has a thin (0.64 mm) Inconel beam entrance window, and  contains 6 \enquote{energy positions} for targets, formed by  5 slabs of 6061 aluminum alloy (previously characterized) which serve as proton energy degraders  between energy positions.
% At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor --- after end-of-bombardment (EoB), decay radiation emitted from these activated stainless steel foils were used to develop radiochromic film (Gafchromic EBT3), revealing the spatial profile of the beam entering and exiting the stack.
After loading all targets in the stack, the lid of the target box is sealed in place, using an inset o-ring to create a water-tight seal, and the box is lowered through a hot cell into the beamline, where it sits electrically isolated.
The specifications of the target stack design for this work is presented in \autoref{tab:stack_table}.

This target stack was assembled and irradiated at the Isotope Production Facility (IPF) at the Los Alamos National Laboratory (LANL), using the LANSCE linear accelerator. 
The stack was irradiated for approximately 2 hours with a nominal current of 1 mA, using a 50 \mmicro s pulse at a frequency of 2 Hz, for an anticipated integral current of 205.9 nAh.
The beam current, measured using an inductive pickup, remained stable under these conditions for the duration of the irradiation, with the exception of approximately 70\,s of downtime, which occurred approximately 3 minutes into irradiation.
The proton beam incident upon the stack's Inconel beam entrance window had an average energy of 100 MeV determined via time-of-flight, with an approximately Gaussian energy distribution width of 0.1 MeV --- this energy profile was used for all later analysis.
At the end of the irradiation, the target stack was withdrawn from the beamline into the IPF hot cell, where it was disassembled and the activated foils removed using robotic manipulators.
The activated foils were cleaned of all surface contamination, and transported to a counting lab for gamma spectrometry, which started approximately 6 hours following end-of-bombardment.

@@

\caption{Specifications of the  target stack design in the present work. The proton beam enters the stack upstream of the 249.8 \mmicro m SS profile monitor, and is transported through the stack in the order presented here. The 6061 aluminum degraders have a measured density of approximately 2.80 g/cm$^3$. Their areal densities were determined using the variance minimization techniques described  in this work  and the earlier paper by Graves \etal\ \cite{Graves2016}. At both the front and rear of the target stack's foils, a 316 stainless steel foil is inserted to serve as a beam profile monitor --- after end-of-bombardment (EoB), decay radiation emitted from these activated stainless steel foils were used to develop radiochromic film (Gafchromic EBT3), revealing the spatial profile of the beam entering and exiting the stack.}

@@

SS profile monitor & 249.8 \mmicro m         & 194.56                                      & 0.29                      \\
Al-1               & 25.0 \mmicro m          & 6.52                                        & 0.72                      \\
Cu-1               & 61.3 \mmicro m          & 53.74                                       & 0.15                      \\
Nb-1               & 30.0 \mmicro m          & 23.21                                       & 0.17                      \\

@@

Al-2               & 25.5 \mmicro m          & 6.48                                        & 0.36                      \\
Cu-2               & 61.8 \mmicro m          & 53.85                                       & 0.17                      \\
Nb-2               & 30.8 \mmicro m          & 22.91                                       & 0.17                      \\

@@

Al-3               & 25.8 \mmicro m          & 6.47                                        & 0.31                      \\
Cu-3               & 61.5 \mmicro m          & 53.98                                       & 0.11                      \\
Nb-3               & 31.0 \mmicro m          & 22.91                                       & 0.24                      \\

@@

Al-4               & 26.3 \mmicro m          & 6.51                                        & 0.41                      \\
Cu-4               & 61.3 \mmicro m          & 53.46                                       & 0.22                      \\
Nb-4               & 30.8 \mmicro m          & 22.55                                       & 0.25                      \\

@@

Al-5               & 26.5 \mmicro m          & 6.48                                        & 0.29                      \\
Cu-5               & 61.5 \mmicro m          & 53.57                                       & 0.11                      \\
Nb-5               & 30.8 \mmicro m          & 22.11                                       & 0.25                      \\

@@

Al-6               & 26.3 \mmicro m          & 6.48                                        & 0.62                      \\
Cu-6               & 62.0 \mmicro m          & 53.84                                       & 0.32                      \\
Nb-6               & 31.3 \mmicro m          & 22.12                                       & 0.13                      \\
SS profile monitor & 124.4 \mmicro m         & 101.34                                      & 0.23                      \\ \bottomrule

@@

\resizebox{\textwidth}{!}{%

@@

}

@@

Notably, this work is the most well-characterized measurement of the \ce{^{nat}Nb}(p,x)\ce{^{90}Mo} reaction below 100 MeV to date, with cross sections measured  at the 4--6\% uncertainty level.
This is important, as it presents the first step towards characterizing this reaction for use as a proton monitor reaction standard below 100 MeV.
% The quantification of \ce{^{90}Mo} activity proved to be one of the most straightforward of all reaction products observed in this work.
\ce{^{nat}Nb}(p,x)\ce{^{90}Mo} can only be populated through the (p,4n) reaction channel, so no corrections for (n,x) contamination channels or decay down the A=90 isobar are needed.
% , which makes activity quantification trivial.
% In addition, 
\ce{^{90}Mo}  possesses seven strong, distinct gamma lines which can easily  be used for its identification and quantification.
Finally, the production of \ce{^{90}Mo}  in the 40--90 MeV region is quite strong, with a peak cross section of approximately 120 mb.
Combining the reaction yield and gamma abundance, the use of approximately 23 mg/cm$^2$ Nb targets easily provided sufficient counting statistics for activity quantification in the 40--90 MeV region.
This result presents the first step towards the use of \ce{^{90}Mo} as a clean and precise   charged particle monitor reaction standard in irradiations up to approximately 24 hours in duration.
